---
title: "TAD Assignment_ Group 6"
author: "NGUYEN Ngoc Nhu Y (22167449)"
date: "2026-03-30"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DATASET INFORMATION

Number of employees: 25

Departments: Operations, Marketing, Engineering, Finance, Sales, HR

Job levels: Mid, Senior, Entry, Executive

Age range: 22 - 53 years

Salary range: \$43,095 - \$322,405

Average performance score: 2.8/5

Average satisfaction score: 6.6/10

# ASSIGNMENT QUESTIONS

# \textcolor{blue}{Question 1: Conduct a thorough exploratory data analysis of the HR dataset.}

## a) Data Structure and Overview

```{r}
# Load dataset
data <- read.csv("group6.csv")
# Structure of dataset
str(data)
head(data)
```

After loading the dataset, an initial overview was conducted to understand its structure and the types of variables present.

The dataset contains **173 observations and 18 variables**, where each observation represents an individual employee.

The variables in the dataset can be classified into two main types:

-   **Quantitative variables**: Age, YearsOfExperience, Salary, PerformanceScore, SatisfactionScore, OvertimeHours, TrainingHours, ProjectsCompleted, TeamSize, RemoteWorkDays, AttendanceRate, and PromotionCount. (these variables are numerical and suitable for statistical analysis)

-   **Qualitative variables:** EmployeeID, Department, Gender, JobLevel, EducationLevel, and OfficeLocation (these variables represent categories or labels)

Each row in the dataset represents an individual employee, while each column corresponds to a specific attribute describing that employee. This structure allows for analysis of workforce characteristics, including performance, compensation, and workplace behaviour.

## b) Compute and present summary statistics

```{r}
# summary statistics for all numeric variables in the dataset
numeric_vars <- sapply(data, is.numeric)

summary(data[, numeric_vars])
sapply(data[, numeric_vars], sd)
```

\newpage

From the output, **Salary shows the strongest evidence of potential outliers:**

-   Q1: 54,284

-   Median: 78,652

-   Q3: 132,833

-   Max: 302,678

-   Mean: 95,076

-   Upper bound = Q3 + 1.5 × IQR = 250,656 (*Using the 1.5×IQR rule, IQR = Q3 − Q1 = 78,549)*

**Since:**

-   **Max: 302,678 \> Upper bound: 250,656**

**-\> this provides clear evidence of the presence of outliers in Salary.**

-   **Mean: 95,076 \> Median: 78,652,**

**-\> further indicating a right-skewed distribution, which is consistent with the presence of high-value outliers.**

Other numeric variables were also examined using the same approach, and no significant outliers were detected based on the IQR rule. Detailed calculations are provided in the Appendix A.

```{r}
#Visualizations
hist(data$Salary, breaks=20, col="lightblue", main ="", xlab="Salary")
```

The histogram of Salary shows a clearly right-skewed distribution.

Most observations are concentrated on the lower end of the scale, approximately between 40,000 and 100,000, where the bars have the highest frequencies. As salary increases, the frequency of observations declines noticeably, as indicated by the decreasing height of the bars. A small number of observations extend far to the right, reaching values close to 300,000, forming a long tail.

This indicates a non-symmetric distribution, where most employees earn lower to mid-range salaries, while a small number receive substantially higher salaries.

```{r}
par(mfrow = c(1, 3))

boxplot(data$Salary,
        col="lightblue",
        main="Salary",
        pch=16)

boxplot(data$YearsOfExperience,
        col="lightgreen",
        main="Experience",
        pch=16)

boxplot(data$Age,
        col="lightpink",
        main="Age",
        pch=16)
```

The boxplots of Salary, Experience, and Age were selected for comparison as they exhibit wider ranges and greater variability relative to other numeric variables, making them more suitable for identifying unusual observations.

Among the three variables, Salary shows the clearest evidence of unusual observations. Specifically, the boxplot displays several data points positioned beyond the upper whisker, indicating the presence of unusually high values. In addition, there is a noticeable gap between the upper edge of the box (IQR) and these points, suggesting a high level of variability at the upper end of the distribution.

In contrast, Experience and Age exhibit more stable and balanced distributions. The boxes (IQRs) are relatively symmetric, with the median located near the center, and all observations fall within the whiskers. No points extend beyond the whiskers.

## c) Key numeric variables

Key numeric variables such as Salary, Experience, and Age were selected as they reflect fundamental workforce characteristics and display clear distributional patterns, making them suitable for analysing trends and relationships within the dataset.

```{r, fig.width=6, fig.height=10}
# Histogram
par(mfrow = c(3,1))

hist(data$Salary,
     col="lightblue",
     main="Distribution of Salary",
     xlab="Salary")

hist(data$YearsOfExperience,
     col="lightgreen",
     main="Distribution of Years of Experience",
     xlab="Years of Experience")

hist(data$Age,
     col="lightpink",
     main="Distribution of Age",
     xlab="Age")
```

**Distribution of Key Variables**

The distribution of Salary is clearly right-skewed, with most observations concentrated at lower to mid salary levels and frequencies decreasing as salary increases. A small number of observations extend toward higher values, forming a long right tail.

The distribution of Experience shows that most employees have relatively low to moderate experience, with frequencies gradually decreasing as experience increases, indicating a mildly right-skewed pattern.

In contrast, Age appears to be more symmetrically distributed, with values relatively evenly spread across the range.

```{r}
#Correlations
cor(data$Salary, data$YearsOfExperience)
cor(data$Salary, data$Age)
cor(data$YearsOfExperience, data$Age)
```

**Correlation Analysis**

There is a strong positive correlation between Salary and Experience (r = 0.869), indicating that employees with more experience tend to earn higher salaries.

Similarly, Salary and Age are also strongly positively correlated (r = 0.861), suggesting that older employees generally earn higher salaries.

In addition, Experience and Age exhibit an extremely strong positive correlation (r = 0.983), indicating that these two variables are highly closely related.

\newpage

## d) Pattern and relationships in the data

```{r}
numeric_vars <- data[, sapply(data, is.numeric)]
cor_matrix <- cor(numeric_vars)

heatmap(cor_matrix, 
        margins = c(8, 8),
        cexRow = 0.8, 
        cexCol = 0.8)
```

A correlation heatmap was first used to provide an overall view of relationships among the numeric variables.

The heatmap reveals several strong positive correlations, most notably among Salary, Years of Experience, and Age. These relationships stand out clearly relative to other variable pairs, suggesting that compensation is closely linked to employees’ tenure and age-related factors. In contrast, many other numeric variables, such as TrainingHours, OvertimeHours, exhibit weaker or less consistent correlations, indicating more independent variation.

\newpage

```{r}
pairs(data[, c("Salary", "YearsOfExperience", "Age")],
      main = "Pairwise Relationships", 
      pch = 16)
```

Guided by the overall patterns, further exploration was conducted using pairwise scatter plots focusing on Salary, Years of Experience, and Age.

The pair plots confirm strong positive linear relationships across all three variables. Salary increases steadily with Years of Experience, indicating that employees with more accumulated experience tend to earn substantially higher salaries. A similar upward trend is observed between Salary and Age, while Years of Experience and Age are almost perfectly linearly related. These visual patterns reinforce the correlation results and suggest that these variables capturing employee seniority within the organisation.

\newpage

```{r}
boxplot(Salary ~ JobLevel, data=data,
        horizontal=TRUE,
        main="Salary by Job Level",
        xlab="Job Level",
        ylab="Salary",
        pch=16)
```

In addition, Job Level was included to explore how salary varies across different employee groups, providing further insight into organisational structure and compensation patterns.

The boxplot reveals clear differences in compensation across job levels. Entry-level employees have lower and less variable salaries, while higher job levels exhibit higher median salaries and greater dispersion, including extreme values. Overall, these patterns suggest that Salary is strongly influenced by experience-related factors and organizational hierarchy.

\newpage

## e) Summary of workforce characteristics

Overall, this workforce is characterized by a moderately wide age range and varying levels of experience, reflecting a mix of early-career and senior employees. Most employees fall into entry- and mid-level job positions, while senior and executive roles represent a smaller proportion of the organization.

Salary displays a strong right-skewed distribution, indicating that while most employees earn low to mid-range salaries, a small number of senior and executive employees receive substantially higher compensation. This pattern aligns with the observed structure of job levels and organizational hierarchy. In contrast, Age and Years of Experience show more balanced distributions and exhibit a very strong positive relationship, reflecting typical career progression patterns.

Exploratory analysis reveals that salary is strongly associated with seniority-related factors, particularly Years of Experience, Age, and Job Level. Visualizations consistently show higher salaries for employees with greater experience and higher job levels. Other workforce-related variables, such as training hours, overtime, and satisfaction, exhibit greater variability and weaker relationships with compensation.

In summary, the dataset describes an organization where compensation is primarily driven by experience and job hierarchy, with clear differences across job levels and relatively stable demographic patterns. These findings provide a strong foundation for subsequent inferential analyses focusing on salary comparisons and workforce structure.

\newpage

# \textcolor{blue}{Question 2: Comparing Salary Across Job Levels}

## a.1) Compare Entry Level vs Senior Level

```{r}
table(data$JobLevel)
aggregate(Salary ~ JobLevel, data, mean)
```

Entry and Senior groups were selected to examine differences in mean salary across career stages, because:

Substantial difference in mean salary:

-   Entry: 51,436

-   Senior: 152,585

-\> The large difference suggests a strong potential for statistically significant variation between the two groups

Sufficient sample sizes:

-   Entry: 70

-   Senior: 44

-\> Both groups meet the requirements for reliable two-sample statistical testing

Moreover, the comparison between Entry and Senior groups is meaningful from an organisational perspective, as it represents early career versus more experienced employees, allowing for an evaluation of how compensation evolves with career progression.

This selection is also consistent with the earlier exploratory analysis, which indicated that salary is strongly associated with experience and job level.

## a.2) Hypotheses testing

The statistical hypotheses are formulated as follows:

-   **Null hypothesis (H0):** The mean salary of Entry-level employees is **equal** to that of Senior-level employees ($μEntry = μSenior$)

-   **Alternative hypothesis (H1):** The mean salary of Entry-level employees is **not equal** to that of Senior-level employees ($μEntry ≠ μSenior$)

\newpage

## b) Welch two sample t-test

```{r}
t.test(Salary ~ JobLevel, 
       data = subset(data, JobLevel %in% c("Entry", "Senior")), 
       var.equal = FALSE)
```

**Results:**

-   **t = -27.085**

-   **df = 55.941**

-   **p-value \< 2.2e - 16**

\newpage

## c) 84% Confidence interval and Effect size

```{r}
# 84% Confident interval
t.test(Salary ~ JobLevel,
       data = subset(data, JobLevel %in% c("Entry", "Senior")),
       var.equal = FALSE,
       conf.level = 0.84)
```

Based on the Welch two-sample t-test results from the previous section, an 84% confidence interval for the difference in mean salaries between Entry-level and Senior-level employees was constructed.

**The resulting confidence interval is: [-106,466.63, -95,830.86]**

This interval does not include zero, indicating that there is a statistically meaningful difference in mean salary between the two groups. Specifically, the negative interval suggests that the mean salary of Entry-level employees is significantly lower than that of Senior employees.

```{r}
# Effect size magnitude
library(effsize)

cohen.d(Salary ~ JobLevel,
        data = subset(data, JobLevel %in% c("Entry", "Senior")))
```

**The effect size: d = -6.014 (large effect)**

The effect size (Cohen’s d = -6.014) is considered extremely large, far exceeding the conventional threshold for a large effect (d = 0.8).

This indicates that the difference in salary between Entry-level and Senior employees is substantial, with minimal overlap between the two distributions. In practical terms, this suggests that job level has a very strong impact on employee compensation.

\newpage

## d) Visual presentations

```{r}
# Boxplot
boxplot(Salary ~ JobLevel,
        data = subset(data, JobLevel %in% c("Entry","Senior")),
        col = c("orange","purple"),
        main = "Salary Comparison: Entry vs Senior",
        ylab = "Salary",
        pch = 16)

## add mean
means <- tapply(data$Salary[data$JobLevel %in% c("Entry","Senior")],
                data$JobLevel[data$JobLevel %in% c("Entry","Senior")],
                mean)

points(1:2, means, col="red", pch=18, cex=1.5)
```

```{r}
# Histogram
par(mfrow = c(1,2))

hist(data$Salary[data$JobLevel=="Entry"],
     col="orange",
     main="Salary of Entry Group",
     xlab="Salary")

hist(data$Salary[data$JobLevel=="Senior"],
     col="purple",
     main="Salary of Senior Group",
     xlab="Salary")
```

To facilitate comparison between Entry-level and Senior employees, multiple visualisations were constructed.

The boxplot with mean values shows that Senior employees have substantially higher median and mean salaries compared to Entry-level employees. In addition, the spread of salaries is wider for the Senior group, indicating greater variability in compensation at higher job levels.

The histograms further illustrate the distribution of salaries within each group. Entry-level salaries are concentrated within a lower range, while Senior-level salaries are distributed at significantly higher values. There is minimal overlap between the two distributions, suggesting a clear separation between the groups.

Overall, these visualisations reinforce the statistical findings from the hypothesis test and confidence interval, confirming a strong and meaningful difference in salary between Entry-level and Senior employees.

\newpage

## d) Conclusion at alpha = 0.02

-   At a significance level of alpha = 0.02, the p-value \< 2.2e - 16, which is much smaller than alpha

**=\> Reject H0**

**=\> There is a statistically significant difference in mean salary between Entry-level and Senior-level employees.**

This conclusion is further supported by the 84% confidence interval for the mean difference, which does not include zero, as well as the very large effect size. These results indicate that the observed salary difference is not only statistically significant but also substantial and consistent in practical terms. The visualizations in part (d) reinforce this finding by showing a clear separation between the salary distributions of the two groups.

From a human resources perspective, a large salary gap between Entry-level and Senior-level employees is expected and appropriate. Senior-level employees typically possess significantly more experience, take on greater responsibilities, and contribute higher strategic value to the organization. As a result, a markedly higher level of compensation reflects a logical and structured pay progression aligned with career advancement.

Overall, the findings suggest that the organization’s compensation strategy clearly reflects job hierarchy and seniority, providing strong financial incentives for employee development and progression. The salary structure appears to be reasonable, transparent, and consistent with standard compensation practices, supporting both organizational effectiveness and long-term workforce motivation.

\newpage

# \textcolor{blue}{Question 3: Analyzing Categorical Variable Relationships}

