---
title: "TAD Assignment_ Group 6"
author: "NGUYEN Ngoc Nhu Y (22167449)"
date: "2026-03-30"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DATASET INFORMATION

Number of employees: 25

Departments: Operations, Marketing, Engineering, Finance, Sales, HR

Job levels: Mid, Senior, Entry, Executive

Age range: 22 - 53 years

Salary range: \$43,095 - \$322,405

Average performance score: 2.8/5

Average satisfaction score: 6.6/10

# ASSIGNMENT QUESTIONS

# \textcolor{blue}{Question 1: Conduct a thorough exploratory data analysis of the HR dataset.}

## a) Data Structure and Overview

```{r}
# Load dataset
data <- read.csv("group6.csv")
# Structure of dataset
str(data)
head(data)
```

After loading the dataset, an initial overview was conducted to understand its structure and the types of variables present.

The dataset contains **173 observations and 18 variables**, where each observation represents an individual employee.

The variables in the dataset can be classified into two main types:

-   **Quantitative variables**: Age, YearsOfExperience, Salary, PerformanceScore, SatisfactionScore, OvertimeHours, TrainingHours, ProjectsCompleted, TeamSize, RemoteWorkDays, AttendanceRate, and PromotionCount. (these variables are numerical and suitable for statistical analysis)

-   **Qualitative variables:** EmployeeID, Department, Gender, JobLevel, EducationLevel, and OfficeLocation (these variables represent categories or labels)

Each row in the dataset represents an individual employee, while each column corresponds to a specific attribute describing that employee. This structure allows for analysis of workforce characteristics, including performance, compensation, and workplace behaviour.

## b) Compute and present summary statistics

```{r}
# summary statistics for all numeric variables in the dataset
numeric_vars <- sapply(data, is.numeric)

summary(data[, numeric_vars])
sapply(data[, numeric_vars], sd)
```

\newpage

From the output, **Salary shows the strongest evidence of potential outliers:**

-   Q1: 54,284

-   Median: 78,652

-   Q3: 132,833

-   Max: 302,678

-   Mean: 95,076

-   Upper bound = Q3 + 1.5 × IQR = 250,656 (*Using the 1.5×IQR rule, IQR = Q3 − Q1 = 78,549)*

**Since:**

-   **Max: 302,678 \> Upper bound: 250,656**

**-\> this provides clear evidence of the presence of outliers in Salary.**

-   **Mean: 95,076 \> Median: 78,652,**

**-\> further indicating a right-skewed distribution, which is consistent with the presence of high-value outliers.**

Other numeric variables were also examined using the same approach, and no significant outliers were detected based on the IQR rule. Detailed calculations are provided in the Appendix A.



