---
title: "TAD Assignment_ Group 6"
author: "NGUYEN Ngoc Nhu Y (22167449)"
date: "2026-03-30"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DATASET INFORMATION

Number of employees: 25

Departments: Operations, Marketing, Engineering, Finance, Sales, HR

Job levels: Mid, Senior, Entry, Executive

Age range: 22 - 53 years

Salary range: \$43,095 - \$322,405

Average performance score: 2.8/5

Average satisfaction score: 6.6/10

# ASSIGNMENT QUESTIONS

# \textcolor{blue}{Question 1: Conduct a thorough exploratory data analysis of the HR dataset.}

## a) Data Structure and Overview

```{r}
# Load dataset
data <- read.csv("group6.csv")
# Structure of dataset
str(data)
head(data)
```

After loading the dataset, an initial overview was conducted to understand its structure and the types of variables present.

The dataset contains **173 observations and 18 variables**, where each observation represents an individual employee.

The variables in the dataset can be classified into two main types:

-   **Quantitative variables**: Age, YearsOfExperience, Salary, PerformanceScore, SatisfactionScore, OvertimeHours, TrainingHours, ProjectsCompleted, TeamSize, RemoteWorkDays, AttendanceRate, and PromotionCount. (these variables are numerical and suitable for statistical analysis)

-   **Qualitative variables:** EmployeeID, Department, Gender, JobLevel, EducationLevel, and OfficeLocation (these variables represent categories or labels)

Each row in the dataset represents an individual employee, while each column corresponds to a specific attribute describing that employee. This structure allows for analysis of workforce characteristics, including performance, compensation, and workplace behaviour.

## b) Compute and present summary statistics

```{r}
# summary statistics for all numeric variables in the dataset
numeric_vars <- sapply(data, is.numeric)

summary(data[, numeric_vars])
sapply(data[, numeric_vars], sd)
```

\newpage

From the output, **Salary shows the strongest evidence of potential outliers:**

-   Q1: 54,284

-   Median: 78,652

-   Q3: 132,833

-   Max: 302,678

-   Mean: 95,076

-   Upper bound = Q3 + 1.5 × IQR = 250,656 (*Using the 1.5×IQR rule, IQR = Q3 − Q1 = 78,549)*

**Since:**

-   **Max: 302,678 \> Upper bound: 250,656**

**-\> this provides clear evidence of the presence of outliers in Salary.**

-   **Mean: 95,076 \> Median: 78,652,**

**-\> further indicating a right-skewed distribution, which is consistent with the presence of high-value outliers.**

Other numeric variables were also examined using the same approach, and no significant outliers were detected based on the IQR rule. Detailed calculations are provided in the Appendix A.

```{r}
#Visualizations
hist(data$Salary, breaks=20, col="lightblue", main ="", xlab="Salary")
```

The histogram of Salary shows a clearly right-skewed distribution.

Most observations are concentrated on the lower end of the scale, approximately between 40000 and 100000, where the bars have the highest frequencies. As salary increases, the frequency of observations declines noticeably, as indicated by the decreasing height of the bars. A small number of observations extend far to the right, reaching values close to 300000, forming a long tail.

This indicates a non-symmetric distribution, where most employees earn lower to mid-range salaries, while a small number receive substantially higher salaries.

```{r}
par(mfrow = c(1, 3))

boxplot(data$Salary,
        col="lightblue",
        main="Salary",
        pch=16)

boxplot(data$YearsOfExperience,
        col="lightgreen",
        main="Experience",
        pch=16)

boxplot(data$Age,
        col="lightpink",
        main="Age",
        pch=16)
```

The boxplots of Salary, Experience, and Age were selected for comparison as they exhibit wider ranges and greater variability relative to other numeric variables, making them more suitable for identifying unusual observations.

Among the three variables, Salary shows the clearest evidence of unusual observations. Specifically, the boxplot displays several data points positioned beyond the upper whisker, indicating the presence of unusually high values. In addition, there is a noticeable gap between the upper edge of the box (IQR) and these points, suggesting a high level of variability at the upper end of the distribution.

In contrast, Experience and Age exhibit more stable and balanced distributions. The boxes (IQRs) are relatively symmetric, with the median located near the center, and all observations fall within the whiskers. No points extend beyond the whiskers.

## c) Key numeric variables

Key numeric variables such as Salary, Experience, and Age were selected as they reflect fundamental workforce characteristics and display clear distributional patterns, making them suitable for analysing trends and relationships within the dataset.

```{r}
```{r, fig.width=6, fig.height=10}
par(mfrow = c(3,1))

hist(data$Salary,
     col="lightblue",
     main="Distribution of Salary",
     xlab="Salary")

hist(data$YearsOfExperience,
     col="lightgreen",
     main="Distribution of Years of Experience",
     xlab="Years of Experience")

hist(data$Age,
     col="lightpink",
     main="Distribution of Age",
     xlab="Age")
```

**Distribution of Key Variables**

The distribution of Salary is clearly right-skewed, with most observations concentrated at lower to mid salary levels and frequencies decreasing as salary increases. A small number of observations extend toward higher values, forming a long right tail.

The distribution of Experience shows that most employees have relatively low to moderate experience, with frequencies gradually decreasing as experience increases, indicating a mildly right-skewed pattern.

In contrast, Age appears to be more symmetrically distributed, with values relatively evenly spread across the range.

```{r}
cor(data$Salary, data$YearsOfExperience)
cor(data$Salary, data$Age)
cor(data$YearsOfExperience, data$Age)
```

**Correlation Analysis**

There is a strong positive correlation between Salary and Experience (r = 0.869), indicating that employees with more experience tend to earn higher salaries.

Similarly, Salary and Age are also strongly positively correlated (r = 0.861), suggesting that older employees generally earn higher salaries.

In addition, Experience and Age exhibit an extremely strong positive correlation (r = 0.983), indicating that these two variables are highly closely related.

\newpage

## d) Pattern and relationships in the data

```{r}
numeric_vars <- data[, sapply(data, is.numeric)]
cor_matrix <- cor(numeric_vars)

heatmap(cor_matrix, 
        margins = c(8, 8),
        cexRow = 0.8, 
        cexCol = 0.8)
```
A correlation heatmap was first used to provide an overall view of relationships among the numeric variables. 

The heatmap reveals several strong positive correlations, most notably among Salary, Years of Experience, and Age. These relationships stand out clearly relative to other variable pairs, suggesting that compensation is closely linked to employees’ tenure and age-related factors. In contrast, many other numeric variables, such as TrainingHours, OvertimeHours, exhibit weaker or less consistent correlations, indicating more independent variation.



