---
title: "TAD1_ Group Assignment_ Group 6"
author: "NGUYEN Ngoc Nhu Y"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DATASET INFORMATION

Number of employees: 25

Departments: Operations, Marketing, Engineering, Finance, Sales, HR

Job levels: Mid, Senior, Entry, Executive

Age range: 22 - 53 years

Salary range: \$43,095 - \$322,405

Average performance score: 2.8/5

Average satisfaction score: 6.6/10

# ASSIGNMENT QUESTIONS

# \textcolor{blue}{Question 1: Conduct a thorough exploratory data analysis of the HR dataset.}

## a) Data Structure and Overview

```{r}
# Load dataset
data <- read.csv("group6.csv")
# Structure of dataset
str(data)
head(data)
```

After loading the dataset, an initial overview was conducted to understand its structure and the types of variables present.

The dataset contains **173 observations and 18 variables**, where each observation represents an individual employee.

The variables in the dataset can be classified into two main types:

-   **Quantitative variables**: Age, YearsOfExperience, Salary, PerformanceScore, SatisfactionScore, OvertimeHours, TrainingHours, ProjectsCompleted, TeamSize, RemoteWorkDays, AttendanceRate, and PromotionCount. (these variables are numerical and suitable for statistical analysis)

-   **Qualitative variables:** EmployeeID, Department, Gender, JobLevel, EducationLevel, and OfficeLocation (these variables represent categories or labels)

Each row in the dataset represents an individual employee, while each column corresponds to a specific attribute describing that employee. This structure allows for analysis of workforce characteristics, including performance, compensation, and workplace behaviour.

\newpage

## b.1) Compute and present summary statistics

```{r}
# Summary statistics for all numeric variables in the dataset
numeric_vars <- sapply(data, is.numeric)

summary(data[, numeric_vars])
sapply(data[, numeric_vars], sd)
```

\newpage

From the output, **Salary shows the strongest evidence of potential outliers:**

-   Q1: 54,284

-   Median: 78,652

-   Q3: 132,833

-   Max: 302,678

-   Mean: 95,076

-   Upper bound = Q3 + 1.5 × IQR = 250,656 (*Using the 1.5×IQR rule, IQR = Q3 − Q1 = 78,549)*

**Since:**

-   **Max: 302,678 \> Upper bound: 250,656**

**-\> this provides clear evidence of the presence of outliers in Salary.**

-   **Mean: 95,076 \> Median: 78,652,**

**-\> further indicating a right-skewed distribution, which is consistent with the presence of high-value outliers.**

Other numeric variables were also examined using the same approach, and no significant outliers were detected based on the IQR rule. Detailed calculations are provided in the Appendix A.

## b.2) Visualizations

```{r}
# Visualize the distribution of Salary to assess skewness and identify potential outliers
hist(data$Salary, 
     breaks=20, 
     col="lightblue", 
     main ="", 
     xlab="Salary")
```

The histogram of Salary shows a clearly right-skewed distribution.

Most observations are concentrated on the lower end of the scale, approximately between 40,000 and 100,000, where the bars have the highest frequencies. As salary increases, the frequency of observations declines noticeably, as indicated by the decreasing height of the bars. A small number of observations extend far to the right, reaching values close to 300,000, forming a long tail.

This indicates a non-symmetric distribution, where most employees earn lower to mid-range salaries, while a small number receive substantially higher salaries.

```{r}
# Compare key numeric variables (Salary, Experience, Age)
par(mfrow = c(1, 3))

boxplot(data$Salary,
        col="lightblue",
        main="Salary",
        pch=16)

boxplot(data$YearsOfExperience,
        col="lightgreen",
        main="Experience",
        pch=16)

boxplot(data$Age,
        col="lightpink",
        main="Age",
        pch=16)
```

The boxplots of Salary, Experience, and Age were selected for comparison as they exhibit wider ranges and greater variability relative to other numeric variables, making them more suitable for identifying unusual observations.

Among the three variables, Salary shows the clearest evidence of unusual observations. Specifically, the boxplot displays several data points positioned beyond the upper whisker, indicating the presence of unusually high values. In addition, there is a noticeable gap between the upper edge of the box (IQR) and these points, suggesting a high level of variability at the upper end of the distribution.

In contrast, Experience and Age exhibit more stable and balanced distributions. The boxes (IQRs) are relatively symmetric, with the median located near the center, and all observations fall within the whiskers. No points extend beyond the whiskers.

## c) Histogram for numeric variables

Key numeric variables such as Salary, Experience, and Age were selected as they reflect fundamental workforce characteristics and display clear distributional patterns, making them suitable for analysing trends and relationships within the dataset.

```{r, fig.width=6, fig.height=7.2}
par(mfrow = c(3,1))

hist(data$Salary,
     col="lightblue",
     main="Distribution of Salary",
     xlab="Salary")

hist(data$YearsOfExperience,
     col="lightgreen",
     main="Distribution of Years of Experience",
     xlab="Years of Experience")

hist(data$Age,
     col="lightpink",
     main="Distribution of Age",
     xlab="Age")
```

**Distribution of Key Variables**

The distribution of Salary is clearly right-skewed, with most observations concentrated at lower to mid salary levels and frequencies decreasing as salary increases. A small number of observations extend toward higher values, forming a long right tail.

The distribution of Experience shows that most employees have relatively low to moderate experience, with frequencies gradually decreasing as experience increases, indicating a mildly right-skewed pattern.

In contrast, Age appears to be more symmetrically distributed, with values relatively evenly spread across the range.

```{r}
#Correlations
cor(data$Salary, data$YearsOfExperience)
cor(data$Salary, data$Age)
cor(data$YearsOfExperience, data$Age)
```

**Correlation Analysis**

There is a strong positive correlation between Salary and Experience (r = 0.869), indicating that employees with more experience tend to earn higher salaries.

Similarly, Salary and Age are also strongly positively correlated (r = 0.861), suggesting that older employees generally earn higher salaries.

In addition, Experience and Age exhibit an extremely strong positive correlation (r = 0.983), indicating that these two variables are highly closely related.

\newpage

## d) Pattern and relationships in the data

```{r}
# Heatmap: Big picture of relationships among numeric variables
numeric_vars <- data[, sapply(data, is.numeric)]
cor_matrix <- cor(numeric_vars)

heatmap(cor_matrix, 
        margins = c(8, 8),
        cexRow = 0.8, 
        cexCol = 0.8)
```

A correlation heatmap was first used to provide an overall view of relationships among the numeric variables.

The heatmap reveals several strong positive correlations, most notably among Salary, Years of Experience, and Age. These relationships stand out clearly relative to other variable pairs, suggesting that compensation is closely linked to employees’ tenure and age-related factors. In contrast, many other numeric variables, such as TrainingHours, OvertimeHours, exhibit weaker or less consistent correlations, indicating more independent variation.

\newpage

```{r}
# Pair plot: Deep dive into key variables (Salary, Experience, Age)
pairs(data[, c("Salary", "YearsOfExperience", "Age")],
      main = "Pairwise Relationships", 
      pch = 16)
```

Guided by the overall patterns, further exploration was conducted using pairwise scatter plots focusing on Salary, Years of Experience, and Age.

The pair plots confirm strong positive linear relationships across all three variables. Salary increases steadily with Years of Experience, indicating that employees with more accumulated experience tend to earn substantially higher salaries. A similar upward trend is observed between Salary and Age, while Years of Experience and Age are almost perfectly linearly related. These visual patterns reinforce the correlation results and suggest that these variables capturing employee seniority within the organisation.

\newpage

```{r}
# Boxplot: Compare Salary across Job Levels to examine group differences
boxplot(Salary ~ JobLevel, data=data,
        horizontal=TRUE,
        main="Salary by Job Level",
        xlab="Job Level",
        ylab="Salary",
        pch=16)
```

In addition, Job Level was included to explore how salary varies across different employee groups, providing further insight into organisational structure and compensation patterns.

The boxplot reveals clear differences in compensation across job levels. Entry-level employees have lower and less variable salaries, while higher job levels exhibit higher median salaries and greater dispersion, including extreme values. Overall, these patterns suggest that Salary is strongly influenced by experience-related factors and organizational hierarchy.

\newpage

## e) Summary of workforce characteristics

Overall, this workforce is characterized by a moderately wide age range and varying levels of experience, reflecting a mix of early-career and senior employees. Most employees fall into entry- and mid-level job positions, while senior and executive roles represent a smaller proportion of the organization.

Salary displays a strong right-skewed distribution, indicating that while most employees earn low to mid-range salaries, a small number of senior and executive employees receive substantially higher compensation. This pattern aligns with the observed structure of job levels and organizational hierarchy. In contrast, Age and Years of Experience show more balanced distributions and exhibit a very strong positive relationship, reflecting typical career progression patterns.

Exploratory analysis reveals that salary is strongly associated with seniority-related factors, particularly Years of Experience, Age, and Job Level. Visualizations consistently show higher salaries for employees with greater experience and higher job levels. Other workforce-related variables, such as training hours, overtime, and satisfaction, exhibit greater variability and weaker relationships with compensation.

In summary, the dataset describes an organization where compensation is primarily driven by experience and job hierarchy, with clear differences across job levels and relatively stable demographic patterns. These findings provide a strong foundation for subsequent inferential analyses focusing on salary comparisons and workforce structure.

\newpage

# \textcolor{blue}{Question 2: Comparing Salary Across Job Levels}

## a.1) Compare Entry Level vs Senior Level

```{r}
table(data$JobLevel)
aggregate(Salary ~ JobLevel, data, mean)
```

Entry and Senior groups were selected to examine differences in mean salary across career stages, because:

Substantial difference in mean salary:

-   Entry: 51,436

-   Senior: 152,585

-\> The large difference suggests a strong potential for statistically significant variation between the two groups

Sufficient sample sizes:

-   Entry: 70

-   Senior: 44

-\> Both groups meet the requirements for reliable two-sample statistical testing

Moreover, the comparison between Entry and Senior groups is meaningful from an organisational perspective, as it represents early career versus more experienced employees, allowing for an evaluation of how compensation evolves with career progression.

This selection is also consistent with the earlier exploratory analysis, which indicated that salary is strongly associated with experience and job level.

## a.2) Hypotheses testing

The statistical hypotheses are formulated as follows:

-   **Null hypothesis (H0):** The mean salary of Entry-level employees is **equal** to that of Senior-level employees ($μEntry = μSenior$)

-   **Alternative hypothesis (H1):** The mean salary of Entry-level employees is **not equal** to that of Senior-level employees ($μEntry ≠ μSenior$)

\newpage

## b) Welch two sample t-test

Given the strong right-skewness and unequal variances observed in salary distributions, the Welch two sample t-test was chosen as it does not assume equal variances between groups.

```{r}
t.test(Salary ~ JobLevel, 
       data = subset(data, JobLevel %in% c("Entry", "Senior")), 
       var.equal = FALSE)
```

**Results:**

-   **t = -27.085**

-   **df = 55.941**

-   **p-value \< 2.2e - 16**

\newpage

## c) 84% Confidence interval and Effect size

```{r}
# 84% Confident interval
t.test(Salary ~ JobLevel,
       data = subset(data, JobLevel %in% c("Entry", "Senior")),
       var.equal = FALSE,
       conf.level = 0.84)
```

Based on the Welch two-sample t-test results from the previous section, an 84% confidence interval for the difference in mean salaries between Entry-level and Senior-level employees was constructed.

**The resulting confidence interval is: [-106,466.63, -95,830.86]**

This interval does not include zero, indicating that there is a statistically meaningful difference in mean salary between the two groups. Specifically, the negative interval suggests that the mean salary of Entry-level employees is significantly lower than that of Senior employees.

```{r}
# Effect size magnitude
library(effsize)

cohen.d(Salary ~ JobLevel,
        data = subset(data, JobLevel %in% c("Entry", "Senior")))
```

**The effect size: d = -6.014 (large effect)**

The effect size (Cohen’s d = -6.014) is considered extremely large, far exceeding the conventional threshold for a large effect (d = 0.8).

This indicates that the difference in salary between Entry-level and Senior employees is substantial, with minimal overlap between the two distributions. In practical terms, this suggests that job level has a very strong impact on employee compensation.

\newpage

## d) Visual presentations

```{r}
# Boxplot
boxplot(Salary ~ JobLevel,
        data = subset(data, JobLevel %in% c("Entry","Senior")),
        col = c("#C9B5D4", "#635BA2"),
        main = "Salary Comparison: Entry vs Senior",
        ylab = "Salary",
        pch = 16)

# add mean
means <- tapply(data$Salary[data$JobLevel %in% c("Entry","Senior")],
                data$JobLevel[data$JobLevel %in% c("Entry","Senior")],
                mean)

points(1:2, means, col="red", pch=18, cex=1.5)
```

```{r}
# Histogram
par(mfrow = c(1,2))

hist(data$Salary[data$JobLevel=="Entry"],
     col="#C9B5D4",
     main="Salary of Entry Group",
     xlab="Salary")

hist(data$Salary[data$JobLevel=="Senior"],
     col="#635BA2",
     main="Salary of Senior Group",
     xlab="Salary")
```

To facilitate comparison between Entry-level and Senior employees, multiple visualizations were constructed.

The boxplot with mean values shows that Senior employees have substantially higher median and mean salaries compared to Entry-level employees. In addition, the spread of salaries is wider for the Senior group, indicating greater variability in compensation at higher job levels.

The histograms further illustrate the distribution of salaries within each group. Entry-level salaries are concentrated within a lower range, while Senior-level salaries are distributed at significantly higher values. There is minimal overlap between the two distributions, suggesting a clear separation between the groups.

Overall, these visualizations reinforce the statistical findings from the hypothesis test and confidence interval, confirming a strong and meaningful difference in salary between Entry-level and Senior employees.

\newpage

## d) Conclusion at alpha = 0.02

-   At a significance level of alpha = 0.02, the p-value \< 2.2e - 16, which is much smaller than alpha

**=\> Reject H0**

**=\> There is a statistically significant difference in mean salary between Entry-level and Senior-level employees.**

This conclusion is further supported by the 84% confidence interval for the mean difference, which does not include zero, as well as the very large effect size. These results indicate that the observed salary difference is not only statistically significant but also substantial and consistent in practical terms. The visualizations in part d) reinforce this finding by showing a clear separation between the salary distributions of the two groups.

From a human resources perspective, a large salary gap between Entry-level and Senior-level employees is expected and appropriate. Senior-level employees typically possess significantly more experience, take on greater responsibilities, and contribute higher strategic value to the organization. As a result, a markedly higher level of compensation reflects a logical and structured pay progression aligned with career advancement.

Overall, the findings suggest that the organization’s compensation strategy clearly reflects job hierarchy and seniority, providing strong financial incentives for employee development and progression. The salary structure appears to be reasonable, transparent, and consistent with standard compensation practices, supporting both organizational effectiveness and long-term workforce motivation.

\newpage

# \textcolor{blue}{Question 3: Analyzing Categorical Variable Relationships}

## a.1) Relationship between Gender and JobLevel

```{r}
# Contingency table between Gender and JobLevel
cont_tab <- table(data$Gender, data$JobLevel)
cont_tab

# Display proportions for better interpretation
prop.table(cont_tab)
```

The contingency table shows that both Male and Female employees are distributed across all job levels, with similar proportions in Entry, Mid, and Senior categories. No clear pattern or imbalance in gender distribution is observed. The “Other” category has very few observations, limiting meaningful interpretation.

## a.2) Hypotheses testing

To formally assess whether there is a relationship between these two variables, the following hypotheses are established:

-   **Null hypothesis (H0):** Gender and JobLevel are **independent**.

-   **Alternative hypothesis (H1):** Gender and JobLevel are **not independent**.

## b) Expected frequencies and chi-squared test

### Expected count

```{r}
test <- chisq.test(cont_tab)
test$expected
```

### Assumption

The assumptions of the chi-squared test were examined prior to conducting the test.

The expected frequencies were calculated, and some cells were found to have expected counts less than 5. This violates the assumption required for the chi-squared approximation to be valid.

Therefore, a **Pearson’s chi-squared test with a simulated p-value** was used to obtain a more reliable result.

### Chi-squared test

```{r}
# Chi-squared test
chi_test <- chisq.test(cont_tab, simulate.p.value = TRUE)
chi_test
```

### Results

-   **X-squared = 6.563**

-   **p-value = 0.3493**

-   **df = NA**

*\*The **theoretical degrees of freedom** for the contingency table **is (3−1)(4−1) = 6**; however, inference is based on a simulated p-value, and therefore df is not used in determining statistical significance.*

## c) Visualizations

```{r}
barplot(cont_tab,
        col = c("pink", "lightblue", "gray"),
        main = "Gender Distribution Across Job Levels",
        xlab = "Job Level",
        ylab = "Count")

legend("topright",
       legend = rownames(cont_tab),
       fill = c("pink", "lightblue", "gray"),
       bty = "n")
```

The stacked bar chart shows the distribution of gender across different job levels.

Overall, the proportions of Male and Female employees remain relatively consistent across Entry, Mid, and Senior levels. There is no clear shift in gender composition as job level increases. Although slight variations are observed, such as a marginally higher proportion of females in Senior roles, these differences are not substantial. The Executive level contains very few observations, making it difficult to draw meaningful conclusions for this category.

Overall, the visualization does not suggest a strong association between Gender and Job Level, which is consistent with the results of the chi-squared test.

## d) Conclusion at alpha = 0.02

At a significance level of alpha = 0.02, the p-value (0.3493) is greater than alpha (0.3493 \> 0.02).

**=\> Fail to reject H0**

**=\> There is no statistically significant association between Gender and Job Level.**

This result suggests that the distribution of employees across job levels does not differ meaningfully by gender. In other words, gender does not appear to influence the likelihood of being in a particular job level within the organisation.

This finding is consistent with the visualizations in part c), which showed relatively similar gender proportions across Entry, Mid, and Senior levels, with no clear pattern of imbalance.

From an organisational perspective, this indicates a relatively balanced workforce structure in terms of gender representation across job levels. There is no strong evidence of gender-based disparities in job level distribution, suggesting that opportunities for advancement may be relatively equitable across genders.

# \textcolor{blue}{Question 4: Simple Linear Regression - Predicting Salary From Years Of Experience}

## a) Simple linear regression model

```{r}
model <- lm(Salary ~ YearsOfExperience, data = data)
summary(model)
```

From the linear regression output the estimated coefficients are:

-   Intercept (b0) = 34,806.1

-   Slope (b1) = 5,349.8

Equation:

-   $\hat{y} = b0 + b1x$

Therefore, the estimated regression equation is:

-   $\hat{y} =34,806.1 + 5,349.8x$

-   $\hat{Salary}$ = 34,806.1 + 5,349.8 x Years Of Experience

\newpage

## b) Interpretation of slope and intercept

```{r}
# Scatter plot with regression line
plot(data$YearsOfExperience, data$Salary,
     col = "orange",
     main = "Salary vs Years of Experience",
     xlab = "Years of Experience",
     ylab = "Salary",
     pch = 16)

abline(model, col = "red", lwd = 2)
```

Overall, both coefficients align with standard compensation practices, where salary growth is systematically tied to increasing experience.

### Slope (b1 = 5,349.8)

The slope coefficient of the regression model is estimated at 5,349.8, indicating that, on average, an employee’s salary is expected to increase by approximately \$5,350 for each additional year of experience, holding all else constant. In the context of HR compensation, this result suggests a strong and positive relationship between work experience and salary, reflecting a pay structure where employees are rewarded for accumulated skills, expertise, and tenure within the organization.

### Intercept (b0 = 34,806.1)

The intercept of the model is estimated at 34,806.1, which represents the predicted salary for an employee with zero years of experience. While this value may not correspond to a realistic salary level in practice, since employees typically enter the workforce with some form of experience, it serves as a baseline reference point for the linear regression model. The intercept helps define the starting point of the salary and experience relationship but should not be interpreted as an actual entry-level wage.

## c) Hypothesis testing on the slope

### Hypotheses

-   **Null Hypothesis (H0):** Years of Experience has **no effect** on Salary (b1 = 0)

-   **Alternative Hypothesis (H1):** Years of Experience **has a significant effect** on Salary (b1 $≠$ 0)

### Test results

From the regression output in part a):

-   t-value = 22.92

-   p-value \< 2.2e-16

### Test conclusion

-   At a significance level of alpha = 0.05, the p-value \< 2.2e - 16, which is much smaller than alpha

**=\> Reject H0**

**=\> There is strong statistical evidence that Years of Experience has a significant linear relationship with Salary.**

\newpage

## d) R-squared

From the regression output in part a):

-   R-squared = 0.7544

This indicates that 75.44% of the variation in Salary can be explained by Years Of Experience.

In other words, years of experience is a strong predictor of salary, accounting for a substantial proportion of the variability in employee compensation. However, the remaining 24.56% of variation is likely influenced by other factors.

Overall, this suggests that while experience plays a major role in determining salary, it is not the only factor affecting compensation.

## e.1) Predict the salary for an employee with 5 years of experience

```{r}
predict(model, newdata = data.frame(YearsOfExperience = 5))
```

-   The predicted value is: 61,554.94

This means that an employee with 5 years of experience is expected to earn around \$61,555, based on the fitted linear relationship between experience and salary.

From a compensation perspective, this result highlights the role of experience in determining salary levels. The model suggests a structured pay progression, where employees are rewarded consistently as they gain experience. However, since not all variation in salary is explained by experience, other factors such as job level and performance should also be considered in compensation decisions.

## e.2) Practical implications for compensation strategy

**Experience-based compensation structure:** The model indicates that salary increases consistently with years of experience, suggesting a structured and predictable compensation system aligned with employee tenure.

**Support for career progression:** The positive relationship between experience and salary provides clear financial incentives for employees to remain with the organization and develop their skills over time.

**Recommendation for HR planning:** The organization may benefit from incorporating additional factors such as performance, job level, and education into its compensation model to ensure a more comprehensive and equitable pay structure.

\newpage

# \textcolor{blue}{Question 5: Comparing Performance Score Across Job Levels}

## a) Boxplot

```{r}
boxplot(PerformanceScore ~ JobLevel,
        data = data,
        col = c("#C9E4D6", "#67BF7F", "#00A06B", "#00676B"),
        main = "Performance Score by Job Level",
        xlab = "Job Level",
        ylab = "Performance Score",
        pch = 16)
```

A boxplot was constructed to visualise the distribution of Performance Score across different Job Levels.

The boxplot shows that the median performance score is approximately 3 for all job levels, indicating a consistent central tendency across groups. The interquartile ranges are also similar, with most scores falling between 2 and 4.

While the Entry level appears to have slightly greater variability, the overall spread of performance scores is comparable across all job levels. No substantial differences in distribution or central tendency are observed.

Overall, the visualisation suggests that performance scores are relatively similar across job levels, with no clear pattern of variation.

\newpage

## b.1) Hypotheses Testing

-   **Null Hypothesis (H0):** Mean Performance Score is the same across all Job Levels

-   **Alternative Hypothesis (H1):** At least one Job Level has a different mean Performance Score

## b.2) ANOVA

```{r}
anova_model <- aov(PerformanceScore ~ JobLevel, data = data)
summary(anova_model)
```

### Results

From the ANOVA output:

-   F-statistic = 0.253

-   p-value = 0.859

### Conclusion

At a significance level of alpha = 0.05, the p-value (0.859) is greater than alpha.

=\> Fail to reject the Null Hypothesis

=\> This indicates that there is no statistically significant difference in mean Performance Score across job levels.

## c) Overall ANOVA test result is not statistically significant

In this case, the overall one‑way ANOVA result is not statistically significant (p‑value = 0.859 \> alpha). This indicates that there is insufficient statistical evidence to conclude that mean Performance Scores differ across Job Levels.

```{r}
# Further support statement
aggregate(PerformanceScore ~ JobLevel, data, mean)
```

This conclusion is further supported by the descriptive statistics. The mean Performance Scores across Job Levels are relatively similar: Entry (2.91), Executive (3.20), Mid (3.11), and Senior (2.93). The observed differences in means are small, suggesting that any variation in Performance Scores across Job Levels is likely due to random variation rather than systematic differences between groups.

As a result, there is no strong justification for conducting post‑hoc pairwise comparisons, since post‑hoc tests are designed to identify specific group differences only when the overall ANOVA is statistically significant.

\
**\*Suppose the overall test result is significant**

Theoretically, ANOVA only answers a general question, whether there is at least one difference in mean Performance Scores across Job Levels, but it does not identify which specific job level pairs differ. Therefore, when the ANOVA result is statistically significant (p‑value \< 0.05), a post‑hoc test such as Tukey’s HSD is required to conduct pairwise comparisons while controlling for multiple testing errors.

The R code for Tukey’s HSD post‑hoc test is as follows:

```{r, eval = FALSE}
TukeyHSD(anova_model)
```

# d) Interpretations of findings

At a significance level of alpha = 0.05, the p-value (0.859) is greater than alpha.

=\> Fail to reject the Null Hypothesis

=\> This indicates that there is no statistically significant difference in mean Performance Score across job levels.

From an organizational perspective, this suggests that employee performance is relatively consistent across different job levels. In other words, higher job levels do not necessarily correspond to higher performance scores.

This may indicate that the organization applies a standardized performance evaluation system, where employees are assessed using similar criteria regardless of their job level. It also suggests that promotion to higher job levels may be influenced more by factors such as experience, tenure, or role responsibilities rather than differences in measured performance.

Overall, the findings imply that while job level strongly influences compensation (as seen in earlier analyses), it does not appear to significantly differentiate employee performance. This reflects a structured organization where performance expectations are relatively uniform across different levels.

\newpage

# TACTICS SUPPORT

# \textcolor{blue}{a) AI Reflection}

This assignment was completed based on concepts learned from lectures and tutorials. AI tool (ChatGPT) was used only as a supportive tool to improve presentation, clarify implementation details, and enhance the readability of the outputs.

Specifically, AI was used to:

-   Refine grammar and improve the wording explanations

-   Improve the formatting and readability of visualizations, including adjusting plot sizes, axis labels, and alignment

-   In 1d) generate the correlation heatmap. The pairwise plot may appear unfamiliar, as it was not covered in this subject; however, it was included based on my knowledge from the Analytics Programming subject (Session 10: Visualization)

-   In 2c) using the library (effsize) and implement Cohen’s d

-   In 2d) enhance the boxplot by adding mean values

# \textcolor{blue}{b) Thinking About Data Subject}

Question 1:

-   Module 3: Descriptive Statistics, Randomization

-   Module 7: Correlation test

Question 2:

-   Module 4: T-test, Confidence Intervals, Bootstrapping

Question 3:

-   Module 2: Chi-squared Test

Question 4:

-   Module 8: Simple Linear Regression

# \textcolor{blue}{c) Analytics Programming Subject}

Module 2: Analyses

-   Session 6: Data (Load Data, View Data,..)

-   Session 10: Data Visualization (Wickham and Grolemund; Chapter 3, 28)

# APPENDICES

# \textcolor{blue}{Appendix A: Detailed Outlier Detection Calculations Using the IQR Method}

## Methodology

Outliers were detected using the standard IQR rule:

-   IQR = Q3 − Q1

-   Lower bound = Q1 − 1.5 × IQR

-   Upper bound = Q3 + 1.5 × IQR

In this case, If Max \> Upper =\> Having Outliers

## 1. Salary

-   Q1 = 54,284

-   Q3 = 132,833

IQR = 132,833 − 54,284 = 78,549

Upper bound = 132,833 + 1.5 × 78,549 = 250,656

Lower bound = 54,284 − 1.5 × 78,549 = −63,539

-   Max value = 302,678

=\> Since 302,678 \> 250,656, this indicates the presence of high-value outliers in Salary.

## 2. Other numeric variables (No outliers detected)

### Age

-   IQR = 42 − 28 = 14

-   Upper = 63

-   Max = 59

### YearsOfExperience

-   IQR = 18 − 4 = 14

-   Upper = 39

-   Max = 35

### PerformanceScore

-   IQR = 4 − 2 = 2

-   Upper = 7

-   Max = 5

### SatisfactionScore

-   IQR = 8 − 3 = 5

-   Upper = 15.5

-   Max = 10

### OvertimeHours

-   IQR = 15 − 5 = 10

-   Upper = 30

-   Max = 19

### TrainingHours

-   IQR = 74 − 25 = 49

-   Upper = 147.5

-   Max = 97

### ProjectsCompleted

-   IQR = 14 − 5 = 9

-   Upper = 27.5

-   Max = 19

### TeamSize

-   IQR = 12 − 6 = 6

-   Upper = 21

-   Max = 14

### RemoteWorkDays

-   IQR = 4 − 1 = 3

-   Upper = 8.5

-   Max = 5

### AttendanceRate

-   IQR = 0.958 − 0.881 = 0.077

-   Upper $≈$ 1.073

-   Max = 0.999

### PromotionCount

-   IQR = 5 − 1 = 4

-   Upper = 11

-   Max = 5
